---
title: "DODGE-Tox compounds"
author: "Jeremy Yang"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

# Introduction

Initial analysis of compound membership for DODGE-Tox, using (1) Canonical SMILES or (2) PubChem CIDs as
unique compound identifier.

RTECS-ReproTox: Reproductive and developmental toxicity dataset excerpted from CDC-RTECS, provided by LeadScope.
RTECS = Registry of Toxic Effects of Chemical Substances (https://cdc.gov/niosh/rtecs).

```{r echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE)
base::date() 
```          
             
```{r echo=FALSE, message=FALSE}
library(readr)
library(data.table)
require(grid)
require(futile.logger)
library(VennDiagram)
flog.appender(appender.console(), name='VennDiagram') #VennDiagram log messages to console.

display_venn <- function(x, ...) {
  grid.newpage()
  venn_object <- venn.diagram(x, filename=NULL, ...)
  grid.draw(venn_object)
}
```

# Compound sets by CFDE-RDKit Canonical SMILES

Canonical SMILES are not globally canonical, thus must be canonicalized via the same algorithm, in this case the RDKit-based CFDE Cheminformatics Db and DevEnv.

## Venn diagrams

```{r}
LS_Cansmi <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/DODGETox/LS_Mapping_UMiami_std_can.smi"), "\t", col_names=c("CANSMI", "LS_ID"))
setDT(LS_Cansmi)
LINCS_Cansmi <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/DODGETox/SM_LINCS_10272021_std_can.smi"), "\t", col_names=c("CANSMI", "LCS_ID"))
setDT(LINCS_Cansmi)
LS2DiseaseArea <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/DODGETox/LS_mapping_MSSM.tsv"), "\t", col_names=c("LS_ID", "DiseaseArea"))
setDT(LS2DiseaseArea)
```

### LINCS, ReproTox-ALL

```{r}
X <- list(
	LINCS = LINCS_Cansmi[!is.na(CANSMI), CANSMI], 
	ReproTox = LS_Cansmi[!is.na(CANSMI), CANSMI])
display_venn(X, fill=c("pink", "cyan"), main="RTECS-ReproTox and LINCS chemical datasets by CFDE-RDKit-Cansmi")
```

### ReproTox-Blood, ReproTox-Cardiovascular, ReproTox-CNS

```{r}
LS_Blood <- LS2DiseaseArea[DiseaseArea=="Blood"]
LS_Blood <- merge(LS_Blood, LS_Cansmi, by="LS_ID", all.x=T, all.y=F)
LS_Cardiovascular <- LS2DiseaseArea[DiseaseArea=="CV"]
LS_Cardiovascular <- merge(LS_Cardiovascular, LS_Cansmi, by="LS_ID", all.x=T, all.y=F)
LS_CNS <- LS2DiseaseArea[DiseaseArea=="CNS"]
LS_CNS <- merge(LS_CNS, LS_Cansmi, by="LS_ID", all.x=T, all.y=F)
X <- list(
	ReproTox_Blood = LS_Blood[!is.na(CANSMI), CANSMI],
	ReproTox_Cardiovascular = LS_Cardiovascular[!is.na(CANSMI), CANSMI],
	ReproTox_CNS = LS_CNS[!is.na(CANSMI), CANSMI]
	)
display_venn(X, fill=c("#0073C2FF", "#EFC000FF", "#868686FF"), main="RTECS-ReproTox chemical datasets by CFDE-RDKit-Cansmi")
```

### LINCS, ReproTox-Blood, ReproTox-Cardiovascular, ReproTox-CNS

```{r}
X <- list(
  LINCS = LINCS_Cansmi[!is.na(CANSMI), CANSMI],
	ReproTox_Blood = LS_Blood[!is.na(CANSMI), CANSMI],
	ReproTox_Cardiovascular = LS_Cardiovascular[!is.na(CANSMI), CANSMI],
	ReproTox_CNS = LS_CNS[!is.na(CANSMI), CANSMI]
	)
display_venn(X, fill=c("pink", "#0073C2FF", "#EFC000FF", "#868686FF"), main="LINCS and RTECS-ReproTox chemical datasets by CFDE-RDKit-Cansmi")
```


# Compound sets by PubChem CID.

* SMILES to CIDs via PubChem REST API.
* CAS Registry Numbers mapped to CIDs via PubChem REST API.
* LINCS compounds (almost) all have CIDs.

Why LINCS cpds with no CIDs? (~1%)

## Venn diagrams

```{r}
SM_LINCS <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/DODGETox/SM_LINCS_10272021.tsv"), "\t")
setDT(SM_LINCS)
SM_LINCS_CIDs <- unique(SM_LINCS[!is.na(SM_PubChem_CID), SM_PubChem_CID])
message(sprintf("SM_LINCS with CID: %d / %d (%.1f%%)", 
	nrow(SM_LINCS[!is.na(SM_PubChem_CID)]),
	nrow(SM_LINCS),
	100*nrow(SM_LINCS[!is.na(SM_PubChem_CID)]) / nrow(SM_LINCS)))
```



```{r}
ReproTox_data_Blood <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/DODGETox/ReproTox_data-Blood_PubChem_cas2cid.tsv"), "\t")
setDT(ReproTox_data_Blood)
ReproTox_data_Blood_CIDs <- unique(ReproTox_data_Blood[, CID])

ReproTox_data_Cardiovascular <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/DODGETox/ReproTox_data-Cardiovascular_PubChem_cas2cid.tsv"), "\t")
setDT(ReproTox_data_Cardiovascular)
ReproTox_data_Cardiovascular_CIDs <- unique(ReproTox_data_Cardiovascular[, CID])

ReproTox_data_CNS <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/DODGETox/ReproTox_data-CNS_PubChem_cas2cid.tsv"), "\t")
setDT(ReproTox_data_CNS)
ReproTox_data_CNS_CIDs <- unique(ReproTox_data_CNS[, CID])
```

### LINCS, ReproTox-ALL

```{r}
X <- list(
	LINCS = SM_LINCS_CIDs,
	ReproTox_ALL = unique(c(ReproTox_data_Blood_CIDs, ReproTox_data_Cardiovascular_CIDs, ReproTox_data_CNS_CIDs))
)
display_venn(X, fill = c("pink", "cyan"), main="(1) LINCS and (2) RTECS-ReproTox chemical datasets (combined) by CID")
```

### ReproTox-Blood, ReproTox-Cardiovascular, ReproTox-CNS

```{r}
X <- list(
	ReproTox_Blood = ReproTox_data_Blood_CIDs, 
	ReproTox_Cardiovascular = ReproTox_data_Cardiovascular_CIDs, 
	ReproTox_CNS = ReproTox_data_CNS_CIDs)
display_venn(X, fill=c("#0073C2FF", "#EFC000FF", "#868686FF"), main="RTECS-ReproTox chemical datasets by CID")
```


### LINCS, ReproTox-Blood, ReproTox-Cardiovascular, ReproTox-CNS

```{r}
X <- list(
	ReproTox_Blood = ReproTox_data_Blood_CIDs, 
	ReproTox_Cardiovascular = ReproTox_data_Cardiovascular_CIDs, 
	ReproTox_CNS = ReproTox_data_CNS_CIDs,
	LINCS = SM_LINCS_CIDs)
display_venn(X, fill = c("#0073C2FF", "#EFC000FF", "#868686FF", "pink"), main="LINCS and RTECS-ReproTox chemical datasets by CID")
```
