---
title: "ReproTox compounds"
author: "Jeremy Yang"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

# Introduction

Initial analysis of compound membership for Repro-Tox, using (1) Canonical SMILES or (2) PubChem CIDs as
unique compound identifier.

RTECS-ReproTox: Reproductive and developmental toxicity dataset excerpted from CDC-RTECS, provided by LeadScope.
RTECS = Registry of Toxic Effects of Chemical Substances (https://cdc.gov/niosh/rtecs).

```{r echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE)
base::date() 
```          
             
```{r echo=FALSE, message=FALSE, results="hide"}
library(readr)
library(data.table)
library(rcdk)
require(grid)
require(futile.logger)
library(VennDiagram)
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger") #Avoid annoying logfiles.
```

```{r echo=FALSE, message=FALSE, results="hide"}
display_venn <- function(x, ...) {
  grid.newpage()
  venn_object <- venn.diagram(x, filename=NULL, ...)
  grid.draw(venn_object)
}
```

# Compound sets by CFDE-RDKit Canonical SMILES

Canonical SMILES are not globally canonical, thus must be canonicalized via the same algorithm, in this case the RDKit-based CFDE Cheminformatics Db and DevEnv.

```{r}
LS_Cansmi <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/LS_Mapping_UMiami_std_can.smi"), "\t", col_names=c("CANSMI", "LS_ID"))
setDT(LS_Cansmi)
LINCS_Cansmi <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/SM_LINCS_10272021_std_can.smi"), "\t", col_names=c("CANSMI", "LCS_ID"))
setDT(LINCS_Cansmi)
LS2DiseaseArea <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/LS_mapping_MSSM.tsv"), "\t", col_names=c("LS_ID", "DiseaseArea"))
setDT(LS2DiseaseArea)
```

## LINCS, ReproTox-ALL

```{r}
X <- list(
	LINCS = LINCS_Cansmi[!is.na(CANSMI), CANSMI], 
	ReproTox = LS_Cansmi[!is.na(CANSMI), CANSMI])
display_venn(X, fill=c("pink", "cyan"), main="RTECS-ReproTox and LINCS chemical datasets by CFDE-RDKit-Cansmi")
```

## ReproTox-Blood, ReproTox-CV, ReproTox-CNS

CV = Cardiovascular, CNS = Central Nervous System

```{r}
LS_Blood <- LS2DiseaseArea[DiseaseArea=="Blood"]
LS_Blood <- merge(LS_Blood, LS_Cansmi, by="LS_ID", all.x=T, all.y=F)
LS_CV <- LS2DiseaseArea[DiseaseArea=="CV"]
LS_CV <- merge(LS_CV, LS_Cansmi, by="LS_ID", all.x=T, all.y=F)
LS_CNS <- LS2DiseaseArea[DiseaseArea=="CNS"]
LS_CNS <- merge(LS_CNS, LS_Cansmi, by="LS_ID", all.x=T, all.y=F)
X <- list(
	ReproTox_Blood = LS_Blood[!is.na(CANSMI), CANSMI],
	ReproTox_CV = LS_CV[!is.na(CANSMI), CANSMI],
	ReproTox_CNS = LS_CNS[!is.na(CANSMI), CANSMI]
	)
display_venn(X, fill=c("#0073C2FF", "#EFC000FF", "#868686FF"), main="RTECS-ReproTox chemical datasets by CFDE-RDKit-Cansmi")
```

## LINCS, ReproTox-Blood, ReproTox-CV, ReproTox-CNS

```{r}
X <- list(
  LINCS = LINCS_Cansmi[!is.na(CANSMI), CANSMI],
	ReproTox_Blood = LS_Blood[!is.na(CANSMI), CANSMI],
	ReproTox_CV = LS_CV[!is.na(CANSMI), CANSMI],
	ReproTox_CNS = LS_CNS[!is.na(CANSMI), CANSMI]
	)
display_venn(X, fill=c("pink", "#0073C2FF", "#EFC000FF", "#868686FF"), main="LINCS and RTECS-ReproTox chemical datasets by CFDE-RDKit-Cansmi")
```

## View chemical structures

Using RCDK R package (Chemistry Dev Kit)

```{r}
moldata <- LS_CV[!is.na(CANSMI)]
depictor <- get.depictor(style='cow')
img_w <- 640
img_h <- 400
n_row <- 5
n_col <- 7
moldata <- moldata[sample(nrow(moldata), n_row*n_col)] #random sample
#par(mar=c(0,0,-.5,0)) #bottom, left, top, right
plot(c(0,1e3), c(0,1e3), pch=1, col="white", axes=FALSE, xlab="", ylab="")
title(main=sprintf("ReproTox-CV sample compounds (%d/%d)", nrow(moldata), nrow(LS_CV[!is.na(CANSMI)])), line = -3, sub="Depicted with RCDK", cex.main=2)
plot.window(xlim=c(0, img_w*n_col), ylim=c(0, img_h*n_row), asp=1.0)
for (i in 1:n_row) {
  for (j in 1:n_col) {
    mol <- parse.smiles(moldata[(i-1)*n_row+(j), CANSMI])
    img <- view.image.2d(mol[[1]], depictor=depictor)
    rasimg <- rasterImage(img, xleft=(j-1)*img_w, xright=j*img_w, ybottom=(i-2)*img_h, ytop=(i-1)*img_h)
  }
}
rasimg
```



```{r}
#view.molecule.2d(mols)
#view.table is fragile.  Needs non-NA strings.

#
#view.table(parse.smiles(moldata$CANSMI), moldata)
```


# Compound sets by PubChem CID.

* SMILES to CIDs via PubChem REST API.
* CAS Registry Numbers mapped to CIDs via PubChem REST API.
* LINCS compounds (almost) all have CIDs.

Why LINCS cpds with no CIDs? (~1%)

```{r warning=FALSE}
SM_LINCS <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/SM_LINCS_10272021.tsv"), "\t", col_types = cols(.default = col_character()))
setDT(SM_LINCS)
SM_LINCS_CIDs <- unique(SM_LINCS[!is.na(SM_PubChem_CID), SM_PubChem_CID])
message(sprintf("SM_LINCS with CID: %d / %d (%.1f%%)", 
	nrow(SM_LINCS[!is.na(SM_PubChem_CID)]),
	nrow(SM_LINCS),
	100*nrow(SM_LINCS[!is.na(SM_PubChem_CID)]) / nrow(SM_LINCS)))
```



```{r}
ReproTox_data_Blood <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/ReproTox_data-Blood_PubChem_cas2cid.tsv"), "\t", col_types = cols(.default = col_character()))
setDT(ReproTox_data_Blood)
ReproTox_data_Blood_CIDs <- unique(ReproTox_data_Blood[, CID])

ReproTox_data_CV <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/ReproTox_data-CV_PubChem_cas2cid.tsv"), "\t", col_types = cols(.default = col_character()))
setDT(ReproTox_data_CV)
ReproTox_data_CV_CIDs <- unique(ReproTox_data_CV[, CID])

ReproTox_data_CNS <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/ReproTox_data-CNS_PubChem_cas2cid.tsv"), "\t", col_types = cols(.default = col_character()))
setDT(ReproTox_data_CNS)
ReproTox_data_CNS_CIDs <- unique(ReproTox_data_CNS[, CID])
```

## LINCS, ReproTox-ALL

```{r}
X <- list(
	LINCS = SM_LINCS_CIDs,
	ReproTox_ALL = unique(c(ReproTox_data_Blood_CIDs, ReproTox_data_CV_CIDs, ReproTox_data_CNS_CIDs))
)
display_venn(X, fill = c("pink", "cyan"), main="(1) LINCS and (2) RTECS-ReproTox chemical datasets (combined) by CID")
```

## ReproTox-Blood, ReproTox-CV, ReproTox-CNS

```{r}
X <- list(
	ReproTox_Blood = ReproTox_data_Blood_CIDs, 
	ReproTox_CV = ReproTox_data_CV_CIDs, 
	ReproTox_CNS = ReproTox_data_CNS_CIDs)
display_venn(X, fill=c("#0073C2FF", "#EFC000FF", "#868686FF"), main="RTECS-ReproTox chemical datasets by CID")
```


## LINCS, ReproTox-Blood, ReproTox-CV, ReproTox-CNS

```{r}
X <- list(
	ReproTox_Blood = ReproTox_data_Blood_CIDs, 
	ReproTox_CV = ReproTox_data_CV_CIDs, 
	ReproTox_CNS = ReproTox_data_CNS_CIDs,
	LINCS = SM_LINCS_CIDs)
display_venn(X, fill = c("#0073C2FF", "#EFC000FF", "#868686FF", "pink"), main="LINCS and RTECS-ReproTox chemical datasets by CID")
```

## Merge into combo file for CFChemDb.

```{r}
LS_Pubchem <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/LS_Mapping_PubChem.tsv"), "\t", col_types = cols(.default = col_character()))
setDT(LS_Pubchem)
setnames(LS_Pubchem, new=c("CID", "SMILES", "LS_ID"))
ReproTox_data <- rbindlist(list(unique(ReproTox_data_Blood[, .(CAS_RN=Name, CID=CID, TERM="Blood")]), unique(ReproTox_data_CNS[, .(CAS_RN=Name, CID=CID, TERM="CNS")]), unique(ReproTox_data_CV[, .(CAS_RN=Name, CID=CID, TERM="CV")])))
ReproTox_data <- merge(ReproTox_data, LS_Pubchem, by="CID", all.x=T, all.y=F)
ReproTox_data <- ReproTox_data[!is.na(LS_ID), .(LS_ID, CAS_RN, CID, SMILES, TERM)]
write_delim(ReproTox_data, paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/ReproTox_export.tsv"), delim="\t")
```

