---
title: "ReproTox compounds"
author: "Jeremy Yang"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

# Introduction

Initial analysis of compound membership for Repro-Tox, using canonical SMILES as
unique compound identifier.

RTECS-ReproTox: Reproductive and developmental toxicity dataset excerpted from CDC-RTECS, provided by LeadScope.
RTECS = Registry of Toxic Effects of Chemical Substances (https://cdc.gov/niosh/rtecs).

```{r echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE)
base::date() 
```          
             
```{r echo=FALSE, message=FALSE, results="hide"}
library(readr)
library(data.table)
library(rcdk)
require(grid)
require(futile.logger)
library(VennDiagram)
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger") #Avoid annoying logfiles.
```

```{r echo=FALSE, message=FALSE, results="hide"}
display_venn <- function(x, ...) {
  grid.newpage()
  venn_object <- venn.diagram(x, filename=NULL, ...)
  grid.draw(venn_object)
}
```

# Compound sets by CFDE-RDKit Canonical SMILES

Canonical SMILES are not globally canonical, thus must be canonicalized via the same algorithm, in this case the RDKit-based CFDE Cheminformatics Db and DevEnv.

```{r}
LS_Cansmi <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/LS_Mapping_std_can.smi"), "\t", col_names=c("CANSMI", "LS_ID"))
setDT(LS_Cansmi)
LINCS_Cansmi <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/SM_LINCS_10272021_std_can.smi"), "\t", col_names=c("CANSMI", "LCS_ID"))
setDT(LINCS_Cansmi)
LS2DiseaseArea <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/LS_mapping_MSSM.tsv"), "\t", col_names=c("LS_ID", "DiseaseArea"))
setDT(LS2DiseaseArea)
```

## LINCS, ReproTox-ALL

```{r}
message(sprintf("ReproTox cansmi count: %d", LS_Cansmi[!is.na(CANSMI), uniqueN(CANSMI)]))
message(sprintf("LINCS cansmi count: %d", LINCS_Cansmi[!is.na(CANSMI), uniqueN(CANSMI)]))
message(sprintf("ReproTox_LINCS common cansmi: %d", length(intersect(LS_Cansmi[!is.na(CANSMI), CANSMI], LINCS_Cansmi[!is.na(CANSMI), CANSMI]))))
X <- list(
	LINCS = LINCS_Cansmi[!is.na(CANSMI), CANSMI], 
	ReproTox = LS_Cansmi[!is.na(CANSMI), CANSMI])
display_venn(X, fill=c("pink", "cyan"), main="RTECS-ReproTox and LINCS chemical datasets by CFDE-RDKit-Cansmi")
```

## ReproTox-Blood, ReproTox-CV, ReproTox-CNS

CV = Cardiovascular, CNS = Central Nervous System

```{r}
LS_Blood <- LS2DiseaseArea[DiseaseArea=="Blood"]
LS_Blood <- merge(LS_Blood, LS_Cansmi, by="LS_ID", all.x=T, all.y=F)
LS_CV <- LS2DiseaseArea[DiseaseArea=="CV"]
LS_CV <- merge(LS_CV, LS_Cansmi, by="LS_ID", all.x=T, all.y=F)
LS_CNS <- LS2DiseaseArea[DiseaseArea=="CNS"]
LS_CNS <- merge(LS_CNS, LS_Cansmi, by="LS_ID", all.x=T, all.y=F)
X <- list(
	ReproTox_Blood = LS_Blood[!is.na(CANSMI), CANSMI],
	ReproTox_CV = LS_CV[!is.na(CANSMI), CANSMI],
	ReproTox_CNS = LS_CNS[!is.na(CANSMI), CANSMI]
	)
display_venn(X, fill=c("#0073C2FF", "#EFC000FF", "#868686FF"), main="RTECS-ReproTox chemical datasets by CFDE-RDKit-Cansmi")
```

## LINCS, ReproTox-Blood, ReproTox-CV, ReproTox-CNS

```{r}
X <- list(
  LINCS = LINCS_Cansmi[!is.na(CANSMI), CANSMI],
	ReproTox_Blood = LS_Blood[!is.na(CANSMI), CANSMI],
	ReproTox_CV = LS_CV[!is.na(CANSMI), CANSMI],
	ReproTox_CNS = LS_CNS[!is.na(CANSMI), CANSMI]
	)
display_venn(X, fill=c("pink", "#0073C2FF", "#EFC000FF", "#868686FF"), main="LINCS and RTECS-ReproTox chemical datasets by CFDE-RDKit-Cansmi")
```

## Export ReproTox-LINCS compounds.

```{r}
ReproTox_LINCS <- merge(rbindlist(list(LS_Blood[!is.na(CANSMI)], LS_CV[!is.na(CANSMI)], LS_CNS[!is.na(CANSMI)])), 
                        LINCS_Cansmi[!is.na(CANSMI)], by="CANSMI", all.x=F, all.y=F)
message(sprintf("ReproTox_LINCS: unique CANSMIs: %d; LS_ID: %d; LCS_ID: %d", ReproTox_LINCS[, uniqueN(CANSMI)], ReproTox_LINCS[, uniqueN(LS_ID)], ReproTox_LINCS[, uniqueN(LCS_ID)]))
write_delim(ReproTox_LINCS, paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/ReproTox_LINCS_export.tsv"), delim="\t")
```

## Merge into combo file for CFChemDb.


```{r}
ReproTox_data_Blood <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/ReproTox_data-Blood_PubChem_cas2cid.tsv"), "\t", col_types = cols(.default = col_character()))
setDT(ReproTox_data_Blood)
ReproTox_data_CV <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/ReproTox_data-CV_PubChem_cas2cid.tsv"), "\t", col_types = cols(.default = col_character()))
setDT(ReproTox_data_CV)
ReproTox_data_CNS <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/ReproTox_data-CNS_PubChem_cas2cid.tsv"), "\t", col_types = cols(.default = col_character()))
setDT(ReproTox_data_CNS)
```

Be sure to export standardized SMILES.

```{r}
LS_Pubchem <- read_delim(paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/LS_Mapping_PubChem.tsv"), "\t", col_types = cols(.default = col_character()))
setDT(LS_Pubchem)
setnames(LS_Pubchem, new=c("CID", "SMILES", "LS_ID"))
LS_Pubchem[, SMILES := NULL]
LS_Pubchem <- merge(LS_Pubchem, LS_Cansmi, by="LS_ID")
setnames(LS_Pubchem, new=c("LS_ID", "CID", "SMILES"))
ReproTox_data <- rbindlist(list(unique(ReproTox_data_Blood[, .(CAS_RN=Name, CID=CID, TERM="Blood")]), unique(ReproTox_data_CNS[, .(CAS_RN=Name, CID=CID, TERM="CNS")]), unique(ReproTox_data_CV[, .(CAS_RN=Name, CID=CID, TERM="CV")])))
ReproTox_data <- merge(ReproTox_data, LS_Pubchem, by="CID", all.x=T, all.y=F)
ReproTox_data <- ReproTox_data[!is.na(LS_ID), .(LS_ID, CAS_RN, CID, SMILES, TERM)]
write_delim(ReproTox_data, paste0(Sys.getenv()["HOME"], "/../data/CFDE/ReproTox/ReproTox_export.tsv"), delim="\t")
```


